---
title: "Reanalysis of BELIV Paper Study Data"
author: "Matthew Kay"
date: "July 18, 2016"
output: html_document
---

Based on `simpleviz-reanalysis.Rmd`.

# Setup

## Imports

```{r setup}
library(magrittr)   #pipe syntax: %>%, %<>%, %$%, ...
library(dplyr)
library(ggplot2)
library(metabayes)
library(tidybayes)
library(rstan)
library(rethinking)
import::from(gamlss.dist,
    TF, dTF, pTF, qTF, rTF)     #scaled and shifted t dist
import::from(gamlss.tr, trun.r) #random number gen for truncated dist
```

## GGplot theme

```{r theme}
theme_set(theme_light())
```

## Stan setup

```{r stan}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Load and clean data

First, let's read in the data:

```{r message=FALSE}
sv <- read.csv("data/simplevis.csv") %>%
    transmute(
        p = Value,
        response = Estimate,
        participant = ID,
        vis = Type,
        confidence = ordered(Confidence, levels=c("low","medium","high")),
        response_time = ResponseTime
    )
head(sv)
```


# Explore first

Let's see the shape of people's error. We will particularly be interested in `squarepie`, which I suspect people are reading essentially as two bar charts (one for the tens column and one for the ones column), but let's start by looking at all of the responses versus the actual probability. We'll include the line `y = x` (black line) and a plain-old linear regression (red) to get a sense of the shape of things:

```{r}
sv %>%
    ggplot(aes(x = p, y=response)) +
    geom_point(alpha = 0.2, size = 1) +
    geom_abline(intercept = 0, slope=1, size = .5) + 
    stat_smooth(method = lm, color="red", se=FALSE) +
    facet_wrap(~vis)
```

Overall, the first impression here is that people's estimates look pretty good in all of the conditions, and that generally speaking there doesn't seem to be any big biases here: the estimates generally fall the line `y = x`. The top end of some of the conditions *might* show some underestimation, but it is not clear at this stage if that is a result of bias or variance: when you have *any* variance in the response up against a ceiling or floor (0 or 100), you should expect the mean to move away from that ceiling/floor. Thus there might be a bias in the *mean* response, but not necessarily the *mode* (most likely) response.

We're also seeing something odd in `squarepie`: there are two faint lines in the error parallel to the central line. Theses lines are at exactly `+-10` from `y = x`. One way to think of a square pie graph might be that it decomposes into two bar graphs: a vertical one representing the 10s column of `p` (0, 10, 20, ...), and a horizontal bar graph stuck on top of it representing the ones column (0, 1, 2, 3, ...). Thus, say the true value (`p`) is `36` and someone misreads the tens column as `4`, they might respond with `45` or `46` or `47`. We can see this phenomenon more easily by looking at density plots of the error (the difference between people's responses and the actual probability depicted):

```{r}
sv %>%
    ggplot(aes(x = response - p)) +
    stat_density() +
    facet_wrap(~vis) +
    geom_vline(xintercept = 0, color="red") +
    coord_cartesian(xlim=c(-20,20))
```

Again we see that all of the visualizations are basically unbiased (the mode of the error---its peak---basically coincides with zero for each vis). Also note the bumps in density at +-10 for `squarepie`, corresponding to off-by-one errors in estimates of the tens column.

# Bias and variance in bars

For convenience, we'll make a version of the `sv` data frame with only entries for the bars vis:

```{r}
sv.bars = sv %>%
    filter(vis == "bars")
```

## A conditional distribution for responses: the discrete truncated Student t

To build a model of people's responses, we'll want some reasonable conditional distribution for a person's responses. That is, given that I know the true value of `p` and which participant is giving me a response, what is a good probability distribution that describes the `response` I should expect them to give me?

For example, we could use a Normal distribution, which would mean that we want a model that takes in `p` and a `participant` and spits out the expected mean and standard deviation of that person's estimates of `p`. However, a straight-up normal distribution isn't the best choice because of ceiling and floor effects. The Normal distribution is defined on `[-Inf, Inf]`, but people's responses cannot be less than 1 or greater than 99. It also isn't that robust to outliers---a few mistakes here or there (or lazy responses from an mturk participant) can drag the mean up or down fairly easily.

Instead, I will use a model that assumes people's estimates follow a discrete truncated Student-*t* distribution. An example of such a distribution is:

```{r}
#make density and random number generator functions for truncated discrete t dist
mu = 2
s = 1.5
nu = 4
truncation_norm = (pTF(99.5, mu, s, nu) - pTF(0.5, mu, s, nu))
d_discrete_t_1to99 = function(x, mu, s, nu) {
    ifelse(
        1 <= x & x <= 99 & x == floor(x), 
        (pTF(x + .5, mu, s, nu) - pTF(x - .5, mu, s, nu)) / truncation_norm,
        0)
}
r_t_.5to99.5 = trun.r(c(0.5, 99.5), "TF", type="both")
r_discrete_t_1to99 = function(...) round(r_t_.5to99.5(...))

#visual explanation of truncated discrete t
mode_label_y = dnorm(mu,mu,s)/truncation_norm*1.1
mean_label_y = dnorm(mu,mu,s)/truncation_norm*1.02
normal_label_y = dnorm(2.85,mu,s)/truncation_norm
t_label_y = dTF(2.9,mu,s,nu)/truncation_norm
discrete_label_y = d_discrete_t_1to99(3, mu, s, nu) * .92
mean. = sum(1:99 %>% {. * d_discrete_t_1to99(., mu, s, nu)})
data_frame(
    x = -5:20,
    p = d_discrete_t_1to99(x, mu, s, nu)
) %>%
    ggplot(aes(x = x)) +
    stat_function(fun = function(x) dTF(x, mu, s, nu) / truncation_norm, 
        geom = "area", color = NA, fill = "gray60", n = 1001) +
    stat_function(fun = function(x) ifelse(x < 0.5, dTF(x, mu, s, nu) / truncation_norm, NA),
        geom = "area", color = NA, fill = "gray85", n = 1001) +
    stat_function(fun = function(x) ifelse(x > 0.5, dnorm(x, mu, s) / truncation_norm, NA),
        color = "gray30", n = 1001) +
    stat_function(fun = function(x) ifelse(x < 0.5, dnorm(x, mu, s) / truncation_norm, NA),
        color = "gray30", linetype="dashed", n = 1001) +
    geom_vline(xintercept = 0) +
    geom_vline(xintercept = 0.5, linetype="dashed") +
    geom_segment(aes(xend = x, y = 0, yend = p), size = 2) +
    geom_vline(xintercept = mu, color="red") +
    geom_vline(xintercept = mean., color="blue", linetype="dashed") +
    annotate("segment", x = mu, y = mode_label_y, xend=5, yend = mode_label_y) +
    annotate("text", x = 5.1, y = mode_label_y, label="Mode", hjust=0) +
    annotate("segment", x = mean., y = mean_label_y, xend=5, yend = mean_label_y) +
    annotate("text", x = 5.1, y = mean_label_y, label="Mean", hjust=0) +
    annotate("segment", x = 2.85, y = normal_label_y, xend=5, yend = normal_label_y) +
    annotate("text", x = 5.1, y = normal_label_y, label="Truncated normal distribution", hjust=0) +
    annotate("segment", x = 2.9, y = t_label_y, xend=5, yend = t_label_y) +
    annotate("text", x = 5.1, y = t_label_y, label="Truncated Student t distribution (nu = 4)", hjust=0) +
    annotate("segment", x = 3, y = discrete_label_y, xend=5, yend = discrete_label_y) +
    annotate("text", x = 5.1, y = discrete_label_y, label="Discrete truncated Student t distribution", hjust=0) +
    xlab("responses") +
    ylab("p(response)")
```

This distribution has the nice property of having its location parameterized by the mean of the corresponding untruncated distribution (which essentially means it is parameterized by the mode), which is useful because we will be hitting ceiling/floor effects since responses in the study could not be < 1 or > 99. At such ceilings/floors, we might *think* people are biased if we examine their mean responses, since the mean will be shifted away from the ceiling/floor, even if their mode response might be unbiased. In the above figure, note how the dotted blue line---the mean---is further from the floor than the solid red line---the mode. The mean shifts further away from the floor the more we increase the variance, even though the mode stays in the same place:

```{r}
for (s in c(1.5, 4, 8, 10)) {
    mean. = sum(1:99 %>% {. * d_discrete_t_1to99(., mu, s, nu)})
    print(data_frame(
        x = -5:20,
        p = d_discrete_t_1to99(x, mu, s, nu)
    ) %>%
        ggplot(aes(x = x)) +
        stat_function(fun = function(x) dTF(x, mu, s, nu) / truncation_norm, 
            geom = "area", color = NA, fill = "gray60", n = 1001) +
        stat_function(fun = function(x) ifelse(x < 0.5, dTF(x, mu, s, nu) / truncation_norm, NA),
            geom = "area", color = NA, fill = "gray85", n = 1001) +
        stat_function(fun = function(x) ifelse(x > 0.5, dnorm(x, mu, s) / truncation_norm, NA),
            color = "gray30", n = 1001) +
        stat_function(fun = function(x) ifelse(x < 0.5, dnorm(x, mu, s) / truncation_norm, NA),
            color = "gray30", linetype="dashed", n = 1001) +
        geom_vline(xintercept = 0) +
        geom_vline(xintercept = 0.5, linetype="dashed") +
        geom_segment(aes(xend = x, y = 0, yend = p), size = 2) +
        geom_vline(xintercept = mu, color="red") +
        geom_vline(xintercept = mean., color="blue", linetype="dashed") +
        xlab("responses") +
        ylab("p(response)")
    )
}
```

As a result, what looks like bias when we examine the mean may actually be variance in responses if we instead consider bias with respect to the mode. Increasingly I have come to believe that mode-unbiasedness is a more useful property to talk about in these kinds of tasks than the more commonly (usually implicit discussed) mean-unbiasedness, because with almost any meaningful variance near a ceiling or floor *any* process will appear to have a biased mean, but may not have a biased mode.

To deal with outliers, I employ a Student t distribution intead of a Normal distribution. This is a common "quick-and-dirty" way to get robust regression: compared to the Normal distribution, the Student t distribution has fatter tails. This means that we *expect* to see more observations off in the tails away from the mode, so when those observations show up they don't drag our mode as far towards them as would happen if we assumed responses were Normally distribution. In fact, we can parameterize tail fatness---that is, learn how likely we should find outliers to be: here, `nu`, sometimes called the *degrees of freedom*, determines how fat tails are. When `nu` is closer to 0, the tails are fatter; as `nu` goes to infinity the t distribution becomes the Normal distribution.

To deal with the fact that responses were recorded as whole numbers, I will use discretization. This approach used could also be thought of as interval censoring at +- .5 around whole numbers, or as measurement error---people's responses were only recorded to the nearest whole number, and this approach accounts for the error introduced by that discretization. Put another way, it assumes that for every whole number `x` from 1 to 99, any answer that would be between `x - 0.5` and `x + 0.5` would be rounded to `x`.

I considered instead using a mode parameterization of a beta distribution, but was unable to get that to fit effectively in Stan when working with mixture models for the squarepie later on. The beta distribution also does not handle fatter tails / outliers (though I considered another modification to the beta distribution it that could handle fatter tails, but also had trouble with fitting in squarepie). Probably these are my limitations and not Stan's, so I will probably return to trying out that sort of model when I get a chance.

## Model specification

Now we'll fit a model with some weakly-informed skeptical priors that assumes the mode response is some linear function of the actual value of `p`, and estimates individuals' variance and the variance of their variance (that is, assumes some people are more or less precise in their estimates than others).

```{r}
m.bars.meta = metastan(
    data = {
        n : int(lower=1)                #number of obervations
        n_participant : int(lower=1)
        
        p : vector[n]
        response : vector[n]
        participant[n] : int
    },
    parameters = {
        mu_intercept : real             #intercept at p = 50 (relative to response = 50)
        mu_b_p : real                   #slope
        
        sigma_intercept : real          #average spread of responses (log-scale)
        sigma_v_participant_raw : vector[n_participant]
        
        sigma_v_participant_sd : real(lower=0)  #variance of participants' variances
        
        #put a lower bound on nu as the first whole number where the t distribution
        #has variance defined
        nu : real(lower=3)
    },
    transformed_parameters = {
        mu : vector[n]
        sigma : vector[n]
        sigma_v_participant : vector[n_participant]

        sigma_v_participant = sigma_v_participant_raw * sigma_v_participant_sd
        for (i in 1:n) {
            sigma[i] <- exp(sigma_intercept + sigma_v_participant[participant[i]])
        }

        #mode response given p. The -50 ... + 50 makes it so that mu_intercept can
        #be interpreted as the offset in the response from 50 when p = 50. It also
        #helps make sampling more efficient by making mu_intercept and m_b_p
        #uncorrelated.
        mu <- mu_intercept + mu_b_p * (p - 50) + 50
    },
    model = {
        #this implies a prior scale of between about .05 and 10
        sigma_intercept ~ normal(-1, 1)
        #prior that most people are within 10x as good or 10x as bad 
        #as the average person's precision
        sigma_v_participant_sd ~ normal(0, 1.15)
        #The next line implies sigma_v_participant ~ normal(0, sigma_v_participant_sd)
        #it is done to improve sampling. See the chapter on optimizing Stan code in the
        #Stan reference manual.
        sigma_v_participant_raw ~ normal(0, 1) 
        
        #prior on bias is around +- 5
        mu_intercept ~ normal(0, 2.5)
        #prior on slope is that it is most likely 1 (this is around 1 +- 1, where a slope of
        #0 would essentially imply "no signal" --- that people's estimates of the response are
        #unrelated to the true p).
        mu_b_p ~ normal(1, 0.5)

        nu ~ exponential(R(1/30))
        
        #the following is the log likelihood for the discrete truncated student t distribution described above
        for (i in 1:n) {
            target %+=% 
                #interval censoring: since responses were recorded only to whole digits,
                #assume a response of x is an interval-censored response from somewhere
                #in [x - 0.5, x + 0.5]
                log_diff_exp(
                    student_t_lcdf(response[i] + 0.5  | nu, mu[i], sigma[i]),
                    student_t_lcdf(response[i] - 0.5  | nu, mu[i], sigma[i])
                ) - 
                #truncation: only responses between 1 and 99 as whole numbers (0.5 and 99.5
                #as real numbers rounded to whole numbers) were allowed
                log_diff_exp(
                    student_t_lcdf(99.5  | nu, mu[i], sigma[i]),
                    student_t_lcdf(0.5  | nu, mu[i], sigma[i])
                )
        }
    }
)
m.bars.stan = stan_model(model_code = as.character(m.bars.meta))
```

```{r}
fit.bars.v = vb(m.bars.stan, data = compose_data(sv.bars), algorithm="meanfield")
```

```{r}
fit.bars = sampling(m.bars.stan, data = compose_data(sv.bars))
```


```{r}
summary(fit.bars, pars=c("mu_intercept","mu_b_p","sigma_intercept","sigma_v_participant_sd", "nu"))
pairs(fit.bars, pars=c("mu_intercept","mu_b_p","sigma_intercept","sigma_v_participant_sd", "nu"))
```


Let's get some fitted values and residuals out from the model so we can inspect it.

```{r}
fitted = fit.bars %>%
    tidy_samples(c(mu, mu_offset, sigma)[i]) %>%
    inner_join(tidy_samples(fit, nu[])) %>%
    mutate(
        predicted = round(r_discrete_t_1to99(n(), mu, sigma, nu))
    ) %>%
    inner_join(mutate(sv.bars, i = 1:n()))

residuals = fitted %>%
    group_by(i, p, response, participant) %>%
    summarise(
        #randomized quantile residuals
        p_resid_max = max(mean(predicted < response + 0.5), 1/n()),
        p_resid_min = min(mean(predicted < response - 0.5), (n() - 1)/n()),
        p_resid = runif(1, p_resid_min, p_resid_max),
        z_resid = qnorm(p_resid),
        mu = mean(mu),
        predicted = sample(predicted, 1)
    ) %>%
    ungroup()
```


```{r}
residuals %>% flatworm(z_resid, participant, ylim=NA) + facet_wrap(~.cuts)
```

```{r}
residuals %>% flatworm(z_resid)
```

```{r}
residuals %>% flatworm(z_resid, p)
```


```{r}
samples = fit.bars %>%
    tidy_samples(c(mu_intercept, mu_b_p, sigma_intercept, sigma_v_participant_sd, nu)[])

predictions = 
    expand.grid(
        p = 1:99
    ) %>%
    group_by(p) %>%
    do({
        cbind(., samples) %>%
            mutate(
                sigma_v_participant = rnorm(n(), 0, sigma_v_participant_sd),
                sigma = exp(sigma_intercept + sigma_v_participant),
                mu = mu_intercept + mu_b_p * (p - 50) + 50,
                predicted = r_discrete_t_1to99(n(), mu, sigma, nu),
                predicted = round(predicted)
            )
    }) %>%
    mutate(error = predicted - p)
```

```{r}
sv.bars %>%
    ggplot(aes(x = response - p)) +
    stat_density(fill=NA, color="red") +
    stat_density(aes(x=predicted - p), fill=NA, color="green", data=residuals, adjust=1)
```

```{r}
fitted10 = fitted %>%
    filter(.sample %in% 15:20)
sv.bars %>%
    ggplot(aes(x = response - p)) +
    stat_density(fill=NA, color="black", size=1.5) +
    stat_density(aes(x=predicted - p, group = .sample), fill=NA, color="blue", data=fitted10, adjust=1) +
    coord_cartesian(xlim=c(-20,20))
```

```{r}
p = predictions %>%
    # select(-mu) %>% rename(mu = predicted) %>%
    mean_qi(mu) %>%
    ggplot(aes(x = p, y = mu)) +
    stat_smooth(aes(y = response), method=lm, data=sv.bars) +
    geom_point(aes(y = response), data=sv.bars, alpha=0.1) +
    geom_ribbon(aes(ymin = mu.lower, ymax = mu.upper), alpha = 0.25) +
    geom_abline(intercept = 0, slope = 1, color="gray50") +
    geom_line(color="red")
p
```

```{r}
p + coord_cartesian(xlim=c(48,52), ylim=c(48,52))
p + coord_cartesian(xlim=c(0,5), ylim=c(0,5))
p + coord_cartesian(xlim=c(95,100), ylim=c(95,100))
```

```{r}
predictions %>%
    filter(.sample %in% 1:10) %>%
    ggplot(aes(x = p, y = predicted)) +
    geom_point(alpha=0.1)
```


```{r}
sv.bars %>%
    ggplot(aes(x = p, y = response)) +
    geom_point(alpha=0.1)
```

```{r}
residuals %>%
    ggplot(aes(x = p, y = predicted)) +
    geom_point(alpha=0.1)
```


# Model for square pie

```{r}
m.squarepie.meta = metastan(
    data = {
        n : int(lower=1)                #number of obervations
        n_participant : int(lower=1)
        
        p : vector[n]
        response : vector[n]
        participant[n] : int
    },
    parameters = {
        mu_intercept : real             #intercept at p = 50 (relative to response = 50)
        mu_b_p : real                   #slope
        
        sigma_intercept : real          #average spread of responses (log-scale)
        sigma_v_participant_raw : vector[n_participant]
        
        sigma_v_participant_sd : real(lower=0)  #variance of participants' variances
        
        theta : real(lower=0, upper=1)  #probability of getting the tens column correct
        
        nu : real(lower=3)
    },
    transformed_parameters = {
        mu : vector[n]
        sigma : vector[n]
        sigma_v_participant : vector[n_participant]
        theta_k : vector[3]

        sigma_v_participant = sigma_v_participant_raw * sigma_v_participant_sd
        for (i in 1:n) {
            sigma[i] <- exp(sigma_intercept + sigma_v_participant[participant[i]])
        }

        theta_k[1] <- (1 - theta) / 2   #probability of off-by-one (below) in tens column
        theta_k[2] <- theta             #probability of getting the tens column exactly correct
        theta_k[3] <- theta_k[1]        #probability of off-by-one (above) in tens column
        
        #mode response given p. The -50 ... + 50 makes it so that mu_intercept can
        #be interpreted as the offset in the response from 50 when p = 50. It also
        #helps make sampling more efficient by making mu_intercept and m_b_p
        #uncorrelated.
        mu <- mu_intercept + mu_b_p * (p - 50) + 50
    },
    model = {
        #this implies a prior scale of between about .05 and 10
        sigma_intercept ~ normal(-0.4, 1.3)
        #prior that most people are within 10x as good or 10x as bad 
        #as the average person's precision
        sigma_v_participant_sd ~ normal(0, 1.15)
        #The next line implies sigma_v_participant ~ normal(0, sigma_v_participant_sd)
        #it is done to improve sampling. See the chapter on optimizing Stan code in the
        #Stan reference manual.
        sigma_v_participant_raw ~ normal(0, 1) 
        
        #prior on bias is around +- 5
        mu_intercept ~ normal(0, 2.5)
        #prior on slope is that it is close to 1 (this is around 1 +- 0.5, which a slope of
        #0 would essentially imply "no signal" --- that people's estimates of the response are
        #unrelated to the true p).
        mu_b_p ~ normal(1, 0.5)

        #prior on the proportion of estimates of the tens column that are correct 
        #(in the neighborhood of 85-100%). Another way to think of this is a prior that
        #the probability of an off-by-one error is about 0-15%.
        theta ~ beta(20,1)
        
        nu ~ exponential(R(1/30))
        
        #the following mess describes a mixture of three censored student-t distributions
        #that make up the likelihood for each observation. The mixture has one distribution
        #centered at mu, one at mu + 10, and one at mu - 10, to account for off-by-one errors
        #in the tens column. I use a student-t distribution with a low df (specifically 3)
        #instead of a normal distribution to account for outliers (essentially as a form of
        #robust regression). The censoring comes in two parts: interval censoring to account
        #for rounding of observations (since the responses are all whole numbers), and 
        #censoring at the endpoints to account for the bounded nature of the responses (from
        #1 to 99).
        for (i in 1:n) {
            log_p[3] : real     #temp for mixture log probabilities
    
            for (k in 1:3) {
                mu_offset : real

                if (k == 1 && mu[i] > 10) {
                    mu_offset = -10
                }
                else if (k == 3 && mu[i] < 90) {
                    mu_offset = 10
                }
                else {
                    mu_offset = 0
                }

                log_p[k] = log_diff_exp(
                    student_t_lcdf(response[i] + 0.5  | nu, mu[i] + mu_offset, sigma[i]),
                    student_t_lcdf(response[i] - 0.5  | nu, mu[i] + mu_offset, sigma[i])
                ) - log_diff_exp(
                    student_t_lcdf(99.5  | nu, mu[i] + mu_offset, sigma[i]),
                    student_t_lcdf(0.5  | nu, mu[i] + mu_offset, sigma[i])
                )

                #theta_k[k] is the probability for this portion of the mixture distribution
                log_p[k] = log_p[k] + log(theta_k[k])
            }
            target %+=% log_sum_exp(log_p)
        }
    },
    generated_quantities = {
        mu_offset : vector[n]
        k : vector[n]

        for (i in 1:n) {
            k[i] <- categorical_rng(theta_k)
            if (k[i] == 1 && mu[i] > 10) {
                mu_offset[i] = -10
            }
            else if (k[i] == 3 && mu[i] < 90) {
                mu_offset[i] = 10
            }
            else {
                mu_offset[i] = 0
            }
        }
    }
)
m.squarepie.stan = stan_model(model_code = as.character(m.meta))
```

```{r}
fit.squarepie = vb(m.stan, data = compose_data(filter(sv, vis == "squarepie")), algorithm="meanfield")
```

```{r}
fit.squarepie = sampling(m.stan, data = compose_data(filter(sv, vis == "squarepie"))
    # , control = list(adapt_delta = 0.9)
    # , iter = 2000 * 6
    # , thin = 6
    )
```

```{r}
summary(fit, pars=c("mu_intercept","mu_b_p","sigma_intercept","sigma_v_participant_sd","nu","theta"))
precis(fit, pars=c("mu_intercept","mu_b_p","theta","sigma_intercept","sigma_v_participant_sd"))
plot(fit, pars=c("mu_intercept","mu_b_p","theta_k2","sigma_intercept","sigma_v_participant_sd"))
```

```{r}
gen.trun(c(0.5,99.5), family="TF", type="both")

samples = tidy_samples(fit, c(
    mu_intercept, mu_b_p, 
    sigma_intercept,
    sigma_v_participant_sd,
    nu
    )[]) %>%
    inner_join(
        tidy_samples(fit, theta_k[..]),
        by = ".sample"
    )


fitted = 
    tidy_samples(fit, c(mu, mu_offset, sigma)[i]) %>%
    inner_join(tidy_samples(fit, nu[])) %>%
    mutate(
        predicted = r_discrete_t_1to99(n(), mu + mu_offset, sigma, nu),
        predicted = round(predicted)
    ) %>%
    inner_join(mutate(filter(sv, vis=="squarepie"), i = 1:n()))

residuals = fitted %>%
    group_by(i, p, response, participant) %>%
    summarise(
        #randomized quantile residuals
        p_resid_max = mean(predicted < response + 0.5),
        p_resid_max = ifelse(p_resid_max == 0, 1/n(), p_resid_max),
        p_resid_min = mean(predicted < response - 0.5),
        p_resid_min = ifelse(p_resid_min == 1, (n() - 1)/n(), p_resid_min),
        p_resid = runif(1, p_resid_min, p_resid_max),
        z_resid = qnorm(p_resid),
        median_predicted = median(predicted),
        predicted = sample(predicted, 1),
        mu = mean(mu)
    ) %>%
    ungroup() %>%
    mutate(
        # z_resid = z_resid / sd(z_resid[is.finite(z_resid)]),
        residual = predicted - response
    )
```

```{r}
residuals %>% flatworm(z_resid, participant, ylim=NA) + facet_wrap(~.cuts)
```

```{r}
residuals %>% flatworm(z_resid)
```

```{r}
residuals %>% flatworm(z_resid, p)
```

```{r}
pairs(fit, pars=c("mu_intercept","mu_b_p","theta","sigma_intercept","sigma_v_participant_sd"))
pairs(fit, pars=c("sigma_intercept","sigma_v_participant_sd","sigma_v_participant[28]","sigma_v_participant[8]","sigma_v_participant[34]"))
pairs(fit, pars=c("sigma_intercept","sigma_v_participant[1]","sigma_v_participant[2]","sigma_v_participant[3]"))
```

```{r}
predictions = 
    expand.grid(
        p = 1:99
    ) %>%
    group_by(p) %>%
    do({
        cbind(., samples) %>%
            group_by(.sample) %>%
            mutate(
                k = rcategorical(1, c(theta_k1, theta_k2, theta_k3))
            ) %>%
            ungroup() %>%
            mutate(
                sigma_v_participant = rnorm(n(), 0, sigma_v_participant_sd),
                sigma = exp(sigma_intercept + sigma_v_participant),
                mu = mu_intercept + mu_b_p * (p - 50) + 50,
                mu_offset = 
                    ifelse(k == 1 & mu > 10, -10,
                    ifelse(k == 3 & mu < 90, 10,
                        0)),
                predicted = r_discrete_t_1to99(n(), mu + mu_offset, sigma, nu),
                predicted = round(predicted)
            )
    }) %>%
    mutate(error = predicted - p)
```

```{r}
predictions %>%
    # filter(p %in% (1:10 * 10)) %>%
    # group_by(p) %>%
    filter(p == 50) %>%
    mutate(
        # mu = mu_intercept + mu_b_p * (p - 50) + 50,
        predicted = r_discrete_t_1to99(n(), mu, sigma, nu)
    ) %>%
    summarise(
        above = mean(predicted > p + 10),
        below = mean(predicted < p - 10),
        theta = 1 - above - below
    )
```


```{r}
sv %>%
    filter(vis == "squarepie") %>%
    ggplot(aes(x = response - p)) +
    stat_density(fill=NA, color="red") +
    stat_density(aes(x=round(predicted) - p), fill=NA, color="green", data=residuals, adjust=1)
```

```{r}
fitted10 = fitted %>%
    filter(.sample %in% 15:20)
sv %>%
    filter(vis == "squarepie") %>%
    ggplot(aes(x = response - p)) +
    stat_density(fill=NA, color="black", size=1.5) +
    stat_density(aes(x=round(predicted) - p, group = .sample), fill=NA, color="blue", data=fitted10, adjust=1) +
    coord_cartesian(xlim=c(-20,20))
```

```{r}
predictions %>%
    select(-mu) %>% rename(mu = predicted) %>%
    mean_qi(mu) %>%
    ggplot(aes(x = p, y = mu)) +
    geom_point(aes(y = response), data=filter(sv, vis=="squarepie"), alpha=0.1) +
    geom_ribbon(aes(ymin = mu.lower, ymax = mu.upper), alpha = 0.25) +
    geom_abline(intercept = 0, slope = 1, color="gray50") +
    geom_line(color="red")
```

```{r}
predictions %>%
    filter(.sample %in% 1:1000) %>%
    ggplot(aes(x = p, y = round(predicted))) +
    geom_point(alpha=0.05)
```

```{r}
residuals %>%
    ggplot(aes(x = p, y = predicted)) +
    geom_point(alpha=0.1)
```
