---
title: "Reanalysis of BELIV Paper Study Data"
author: "Matthew Kay"
date: "July 18, 2016"
output: html_document
---

Based on `simpleviz-reanalysis.Rmd`.

# Setup

## Imports

```{r setup}
library(dplyr)
library(ggplot2)
library(metabayes)
library(tidybayes)
library(rstan)
library(rethinking)
import::from(gamlss, gamlss, random, re)
import::from(gamlss.dist, 
    NO, dNO, pNO, qNO, rNO, 
    TF, dTF, pTF, qTF, rTF)
import::from(gamlss.tr, gen.trun, trun.r)
```

## GGplot theme

```{r theme}
theme_set(theme_light())
```

## Stan setup

```{r stan}
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

# Load and clean data

First, let's read in the data:

```{r message=FALSE}
sv <- read.csv("data/simplevis.csv") %>%
    transmute(
        p = Value,
        response = Estimate,
        participant = ID,
        vis = Type,
        confidence = ordered(Confidence, levels=c("low","medium","high")),
        response_time = ResponseTime
    )
head(sv)
```


# Explore first

Let's see the shape of people's error. We will particularly be interested in `squarepie`, which I suspect people are reading essentially as two bar charts (one for the tens column and one for the ones column), but let's start by looking at all of the responses versus the actual probability. We'll include the line `y = x` (black line) and a plain-old linear regression (red) to get a sense of the shape of things:

```{r}
sv %>%
    ggplot(aes(x = p, y=response)) +
    geom_point(alpha = 0.2, size = 1) +
    geom_abline(intercept = 0, slope=1, size = .5) + 
    stat_smooth(method = lm, color="red", se=FALSE) +
    facet_wrap(~vis)
```

Overall, the first impression here is that people's estimates look pretty good in all of the conditions, and that generally speaking there doesn't seem to be any big biases here. The top end of some of the conditions *might* show some underestimation, but we will need a model that accounts for ceiling and floor effects to really decide if that is the case: when you have *any* variance in the response up against a ceiling or floor (0 or 100), you should expect the mean to move away from that ceiling/floor. Thus there might be a bias in the *mean* response, but not necessarily the *mode* response.

We're also see something odd in `squarepie`: there are two faint lines in the error parallel to the central line. Theses lines are at exactly `+-10` from `y = x`. One way to think of a square pie graph might be that it decomposes into two bar graphs: a vertical one representing the 10s column of `p` (0, 10, 20, ...), and a horizontal bar graph stuck on top of it representing the ones column (0, 1, 2, 3, ...). Thus, say the true value (`p`) is `36` and someone misreads the tens column as `4`, they might respond with `45` or `46` or `47`. We can see this phenomenon more easily by looking at density plots of the error (the difference between people's responses and the actual probability depicted):

```{r}
sv %>%
    ggplot(aes(x = response - p)) +
    stat_density() +
    facet_wrap(~vis) +
    geom_vline(xintercept = 0, color="red") +
    coord_cartesian(xlim=c(-20,20))
```

Again we see that all of the visualizations are basically unbiased. Also note the bumps in density at +-10 for `squarepie`, corresponding to off-by-one errors in estimates of the tens column.

Let's first see if we can do a reasonable job of decomposing bias and variance in one of the charts, bars.


```{r}
sv %>%
    filter(vis == "bars") %>%
    filter(participant %in% paste0("p", 40:48)) %>%
    ggplot(aes(x = p, y=response)) +
    geom_point() +
    # geom_line(aes(color = participant)) +
    stat_smooth(aes(group=participant), method=lm, se=FALSE) +
    geom_abline(intercept = 0, slope=1, linetype="dashed")+
    facet_wrap(~participant)
```

```{r}
sv %>%
    filter(vis == "squarepie") %>%
    filter(participant %in% paste0("p", 10:20)) %>%
    ggplot(aes(x = response - p)) +
    stat_density() + 
    geom_vline(xintercept = 0, linetype="dashed", color="red") +
    geom_vline(xintercept = -10, linetype="dashed", color="red") +
    geom_vline(xintercept = 10, linetype="dashed", color="red") 
    # facet_wrap(~participant)
```

And that is what we see. Let's see that as estimate versus value:

```{r}
sv %>%
    ggplot(aes(x = Value, y = Estimate)) +
    geom_point(alpha = 0.1) +
    geom_abline(intercept = 0, slope = 1, color = "red") +
    facet_wrap(~Type)
```

And again we see the additional lines at +-10 for the square pie. I assume this happens when someone's estimate of the tens column (reading of the vertical "bar"" that is part of the square pie) is off by one, but their estimate of the ones column (reading of the horizontal "bar" that is part of the square pie) is still decent. 

Another thing worth noting: overall, people don't appear to be biased in any of the conditions: note how the estimates fall along the red line for `y = x`.

Let's fit a model to this (ignoring the off-by-ten errors in square pie for now):


# Bias and variance in bars

For convenience, we'll make a version of the `sv` data frame with only entries for the bars vis:

```{r}
sv.bars = sv %>%
    filter(vis == "bars")
```

## Description of conditional distribution of responses

To build a model of people's responses, we'll want some reasonable conditional distribution for a person's responses. That is, given that I know the true value of `p` and which participant is giving me a response, what is a good probability distribution that describes the value of their `response`?

For example, we could use a Normal distribution, which would mean that we want a model that takes in `p` and a `participant` and spits out the expected mean and standard deviation of that person's estimates of `p`. However, a straight-up normal distribution isn't the best choice because of ceiling and floor effects: people's responses could not be less than 1 or greater than 99. It also isn't that robust to outliers.

Instead, I will use a model that assumes people's estimates follow a truncated, discrete Student-*t* distribution. An example of such a distribution is:

```{r}
mu = 2
s = 1.5
nu = 4
truncation_norm = (pTF(99.5, mu, s, nu) - pTF(0.5, mu, s, nu))
mode_label_y = dnorm(mu,mu,s)/truncation_norm*1.1
mean_label_y = dnorm(mu,mu,s)/truncation_norm*1.02
normal_label_y = dnorm(2.85,mu,s)/truncation_norm
t_label_y = dTF(2.9,mu,s,nu)/truncation_norm
d_discrete_t_1to99 = function(x, mu, s, nu) {
    ifelse(
        1 <= x & x <= 99 & x == floor(x), 
        (pTF(x + .5, mu, s, nu) - pTF(x - .5, mu, s, nu)) / truncation_norm,
        0)
}
discrete_label_y = d_discrete_t_1to99(3) * .92
mean. = sum(1:99 %>% {. * d_discrete_t_1to99(., mu, s, nu)})
data_frame(
    x = -5:20,
    p = d_discrete_t_1to99(x, mu, s, nu)
) %>%
    ggplot(aes(x = x)) +
    stat_function(fun = function(x) dTF(x, mu, s, nu) / truncation_norm, 
        geom = "area", color = NA, fill = "gray60", n = 1001) +
    stat_function(fun = function(x) ifelse(x < 0.5, dTF(x, mu, s, nu) / truncation_norm, NA),
        geom = "area", color = NA, fill = "gray85", n = 1001) +
    stat_function(fun = function(x) ifelse(x > 0.5, dnorm(x, mu, s) / truncation_norm, NA),
        color = "gray30", n = 1001) +
    stat_function(fun = function(x) ifelse(x < 0.5, dnorm(x, mu, s) / truncation_norm, NA),
        color = "gray30", linetype="dashed", n = 1001) +
    geom_vline(xintercept = 0) +
    geom_vline(xintercept = 0.5, linetype="dashed") +
    geom_segment(aes(xend = x, y = 0, yend = p), size = 2) +
    geom_vline(xintercept = mu, color="red") +
    geom_vline(xintercept = mean., color="blue", linetype="dashed") +
    annotate("segment", x = mu, y = mode_label_y, xend=5, yend = mode_label_y) +
    annotate("text", x = 5.1, y = mode_label_y, label="Mode", hjust=0) +
    annotate("segment", x = mean., y = mean_label_y, xend=5, yend = mean_label_y) +
    annotate("text", x = 5.1, y = mean_label_y, label="Mean", hjust=0) +
    annotate("segment", x = 2.85, y = normal_label_y, xend=5, yend = normal_label_y) +
    annotate("text", x = 5.1, y = normal_label_y, label="Truncated normal distribution", hjust=0) +
    annotate("segment", x = 2.9, y = t_label_y, xend=5, yend = t_label_y) +
    annotate("text", x = 5.1, y = t_label_y, label="Truncated Student t distribution (nu = 4)", hjust=0) +
    annotate("segment", x = 3, y = discrete_label_y, xend=5, yend = discrete_label_y) +
    annotate("text", x = 5.1, y = discrete_label_y, label="Discrete truncated Student t distribution", hjust=0) +
    xlab("responses") +
    ylab("p(response)")
```

This distribution has the nice property of having its location parameterized by the mean of the corresponding untruncated distribution (which essentially means it is parameterized by the mode), which is useful because we will be hitting ceiling/floor effects since responses in the study could not be < 1 or > 99. At such ceilings/floors, we might *think* people are biased if we examine their mean responses, since the mean will be shifted away from the ceiling/floor, even if their mode response might be unbiased. Thus, it is useful to examine the location of the mode. Increasingly I have come to believe that mode-unbiasedness is a more useful property to talk about in these kinds of tasks than the more commonly (usually implicit discussed) mean-unbiasedness, because with almost any meaningful variance near a ceiling or floor *any* process will appear to have a biased mean, but may not have a biased mode.

Compared to the Normal distribution, the Student t distribution has fatter tails, which helps make it robust to outliers. In fact, we can parameterize tail fatness: here, `nu`, sometimes called the *degrees of freedom*); as `nu` goes to infinity the t distribution becomes the Normal distribution.

The discretization used could also be thought of as interval censoring at +- .5 around whole numbers, or as measurement error---people's responses were only recorded to the nearest whole number, and this approach accounts for the error introduced by that discretization. Put another way, it assumes that for every whole number `x` from 1 to 99, any answer that would be between `x - 0.5` and `x + 0.5` would be rounded to `x`.

I considered instead using a mode parameterization of a beta distribution, but was unable to get that to fit effectively in Stan when working with mixture models for the squarepie later on. The beta distribution also does not handle fatter tails / outliers (though I considered another modification to the beta distribution it that could handle fatter tails, but also had trouble with fitting in squarepie---sometime when I get a chance I may spend more time getting that model right and writing it up).

## Model specification

Now we'll fit a model with some weakly-informed skeptical priors that assumes the mode response is some linear function of the actual value of `p`, and estimates individuals' variance and the variance of their variance (that is, assumes some people are more or less precise in their estimates than others).

```{r}
m.bars.meta = metastan(
    data = {
        n : int(lower=1)                #number of obervations
        n_participant : int(lower=1)
        
        p : vector[n]
        response : vector[n]
        participant[n] : int
    },
    parameters = {
        mu_intercept : real             #intercept at p = 50 (relative to response = 50)
        mu_b_p : real                   #slope
        
        sigma_intercept : real          #average spread of responses (log-scale)
        sigma_v_participant_raw : vector[n_participant]
        
        sigma_v_participant_sd : real(lower=0)  #variance of participants' variances
        
        #put a lower bound on nu as the first whole number where the t distribution
        #has variance defined
        nu : real(lower=3)
    },
    transformed_parameters = {
        mu : vector[n]
        sigma : vector[n]
        sigma_v_participant : vector[n_participant]

        sigma_v_participant = sigma_v_participant_raw * sigma_v_participant_sd
        for (i in 1:n) {
            sigma[i] <- exp(sigma_intercept + sigma_v_participant[participant[i]])
        }

        #mode response given p. The -50 ... + 50 makes it so that mu_intercept can
        #be interpreted as the offset in the response from 50 when p = 50. It also
        #helps make sampling more efficient by making mu_intercept and m_b_p
        #uncorrelated.
        mu <- mu_intercept + mu_b_p * (p - 50) + 50
    },
    model = {
        #this implies a prior scale of between about .05 and 10
        sigma_intercept ~ normal(-1, 1)
        #prior that most people are within 10x as good or 10x as bad 
        #as the average person's precision
        sigma_v_participant_sd ~ normal(0, 1.15)
        #The next line implies sigma_v_participant ~ normal(0, sigma_v_participant_sd)
        #it is done to improve sampling. See the chapter on optimizing Stan code in the
        #Stan reference manual.
        sigma_v_participant_raw ~ normal(0, 1) 
        
        #prior on bias is around +- 5
        mu_intercept ~ normal(0, 2.5)
        #prior on slope is that it is most likely 1 (this is around 1 +- 1, where a slope of
        #0 would essentially imply "no signal" --- that people's estimates of the response are
        #unrelated to the true p).
        mu_b_p ~ normal(1, 0.5)

        nu ~ exponential(R(1/30))
        
        #the following is the log likelihood for the discrete truncated student t distribution described above
        for (i in 1:n) {
            target %+=% 
                #interval censoring: since responses were recorded only to whole digits,
                #assume a response of x is an interval-censored response from somewhere
                #in [x - 0.5, x + 0.5]
                log_diff_exp(
                    student_t_lcdf(response[i] + 0.5  | nu, mu[i], sigma[i]),
                    student_t_lcdf(response[i] - 0.5  | nu, mu[i], sigma[i])
                ) - 
                #truncation: only responses between 1 and 99 as whole numbers (0.5 and 99.5
                #as real numbers rounded to whole numbers) were allowed
                log_diff_exp(
                    student_t_lcdf(99.5  | nu, mu[i], sigma[i]),
                    student_t_lcdf(0.5  | nu, mu[i], sigma[i])
                )
        }
    }
)
m.bars.stan = stan_model(model_code = as.character(m.bars.meta))
```

```{r}
fit.bars.v = vb(m.bars.stan, data = compose_data(sv.bars), algorithm="meanfield")
```

```{r}
fit.bars = sampling(m.bars.stan, data = compose_data(sv.bars))
```


```{r}
summary(fit.bars, pars=c("mu_intercept","mu_b_p","sigma_intercept","sigma_v_participant_sd", "nu"))
pairs(fit.bars, pars=c("mu_intercept","mu_b_p","sigma_intercept","sigma_v_participant_sd", "nu"))
```


Let's get some fitted values and residuals out from the model so we can inspect it.

```{r}
gen.trun(c(0.5,99.5), family="TF", type="both")
pt.trun = function(x, mu, sigma, nu) {
    ifelse(x < 0.5, 0,
    ifelse(x > 99.5, 1,
        (pTF(x, mu, sigma, nu) - pTF(0.5, mu, sigma, nu))  / (pTF(99.5, mu, sigma, nu) - pTF(0.5, mu, sigma, nu))
    ))
}
dt.trun = function(x, mu, sigma, nu) {
    ifelse(x < 0.5 | x > 99.5, 0,
        dTF(x, mu, sigma, nu) / (pTF(99.5, mu, sigma, nu) - pTF(0.5, mu, sigma, nu))
    )
}
invert = function (f, lower = 0.5, upper = 99.5) {
    #invert a function (based on http://stackoverflow.com/a/10081571)
    function (y, ...) uniroot((function (x) f(x, ...) - y), lower = lower, upper = upper)$root
}
qt.trun = invert(pt.trun)
```

```{r}
fitted = fit.bars %>%
    tidy_samples(c(mu, mu_offset, sigma)[i]) %>%
    inner_join(tidy_samples(fit, nu[])) %>%
    mutate(
        predicted = round(rTFtr(n(), mu, sigma, nu))
    ) %>%
    inner_join(mutate(sv.bars, i = 1:n()))

residuals = fitted %>%
    group_by(i, p, response, participant) %>%
    summarise(
        #randomized quantile residuals
        p_resid_max = max(mean(predicted < response + 0.5), 1/n()),
        p_resid_min = min(mean(predicted < response - 0.5), (n() - 1)/n()),
        p_resid = runif(1, p_resid_min, p_resid_max),
        z_resid = qnorm(p_resid),
        mu = mean(mu),
        predicted = sample(predicted, 1)
    ) %>%
    ungroup()
```


```{r}
residuals %>% flatworm(z_resid, participant, ylim=NA) + facet_wrap(~.cuts)
```

```{r}
residuals %>% flatworm(z_resid)
```

```{r}
residuals %>% flatworm(z_resid, p)
```


```{r}
samples = fit.bars %>%
    tidy_samples(c(mu_intercept, mu_b_p, sigma_intercept, sigma_v_participant_sd, nu)[])

predictions = 
    expand.grid(
        p = 1:99
    ) %>%
    group_by(p) %>%
    do({
        cbind(., samples) %>%
            mutate(
                sigma_v_participant = rnorm(n(), 0, sigma_v_participant_sd),
                sigma = exp(sigma_intercept + sigma_v_participant),
                mu = mu_intercept + mu_b_p * (p - 50) + 50,
                predicted = rTFtr(n(), mu, sigma, nu),
                predicted = round(predicted)
            )
    }) %>%
    mutate(error = predicted - p)
```

```{r}
sv.bars %>%
    ggplot(aes(x = response - p)) +
    stat_density(fill=NA, color="red") +
    stat_density(aes(x=predicted - p), fill=NA, color="green", data=residuals, adjust=1)
```

```{r}
fitted10 = fitted %>%
    filter(.sample %in% 15:20)
sv.bars %>%
    ggplot(aes(x = response - p)) +
    stat_density(fill=NA, color="black", size=1.5) +
    stat_density(aes(x=predicted - p, group = .sample), fill=NA, color="blue", data=fitted10, adjust=1) +
    coord_cartesian(xlim=c(-20,20))
```

```{r}
p = predictions %>%
    # select(-mu) %>% rename(mu = predicted) %>%
    mean_qi(mu) %>%
    ggplot(aes(x = p, y = mu)) +
    geom_point(aes(y = response), data=sv.bars, alpha=0.1) +
    geom_ribbon(aes(ymin = mu.lower, ymax = mu.upper), alpha = 0.25) +
    geom_abline(intercept = 0, slope = 1, color="gray50") +
    geom_line(color="red")
p
```

```{r}
p + coord_cartesian(xlim=c(48,52), ylim=c(48,52))
p + coord_cartesian(xlim=c(0,5), ylim=c(0,5))
p + coord_cartesian(xlim=c(95,100), ylim=c(95,100))
```

```{r}
predictions %>%
    filter(.sample %in% 1:10) %>%
    ggplot(aes(x = p, y = predicted)) +
    geom_point(alpha=0.1)
```

```{r}
residuals %>%
    ggplot(aes(x = p, y = predicted)) +
    geom_point(alpha=0.1)
```


# Model for square pie

```{r}
m.meta = metastan(
    data = {
        n : int(lower=1)                #number of obervations
        n_participant : int(lower=1)
        
        p : vector[n]
        response : vector[n]
        participant[n] : int
    },
    parameters = {
        mu_intercept : real             #intercept at p = 50 (relative to response = 50)
        mu_b_p : real                   #slope
        
        sigma_intercept : real          #average spread of responses (log-scale)
        sigma_v_participant_raw : vector[n_participant]
        
        sigma_v_participant_sd : real(lower=0)  #variance of participants' variances
        
        theta : real(lower=0, upper=1)  #probability of getting the tens column correct
        
        nu : real(lower=3)
    },
    transformed_parameters = {
        mu : vector[n]
        sigma : vector[n]
        sigma_v_participant : vector[n_participant]
        theta_k : vector[3]

        sigma_v_participant = sigma_v_participant_raw * sigma_v_participant_sd
        for (i in 1:n) {
            sigma[i] <- exp(sigma_intercept + sigma_v_participant[participant[i]])
        }

        theta_k[1] <- (1 - theta) / 2   #probability of off-by-one (below) in tens column
        theta_k[2] <- theta             #probability of getting the tens column exactly correct
        theta_k[3] <- theta_k[1]        #probability of off-by-one (above) in tens column
        
        #mode response given p. The -50 ... + 50 makes it so that mu_intercept can
        #be interpreted as the offset in the response from 50 when p = 50. It also
        #helps make sampling more efficient by making mu_intercept and m_b_p
        #uncorrelated.
        mu <- mu_intercept + mu_b_p * (p - 50) + 50
    },
    model = {
        #this implies a prior scale of between about .05 and 10
        sigma_intercept ~ normal(-0.4, 1.3)
        #prior that most people are within 10x as good or 10x as bad 
        #as the average person's precision
        sigma_v_participant_sd ~ normal(0, 1.15)
        #The next line implies sigma_v_participant ~ normal(0, sigma_v_participant_sd)
        #it is done to improve sampling. See the chapter on optimizing Stan code in the
        #Stan reference manual.
        sigma_v_participant_raw ~ normal(0, 1) 
        
        #prior on bias is around +- 5
        mu_intercept ~ normal(0, 2.5)
        #prior on slope is that it is close to 1 (this is around 1 +- 0.5, which a slope of
        #0 would essentially imply "no signal" --- that people's estimates of the response are
        #unrelated to the true p).
        mu_b_p ~ normal(1, 0.5)

        #prior on the proportion of estimates of the tens column that are correct 
        #(in the neighborhood of 85-100%). Another way to think of this is a prior that
        #the probability of an off-by-one error is about 0-15%.
        theta ~ beta(20,1)
        
        nu ~ exponential(R(1/30))
        
        #the following mess describes a mixture of three censored student-t distributions
        #that make up the likelihood for each observation. The mixture has one distribution
        #centered at mu, one at mu + 10, and one at mu - 10, to account for off-by-one errors
        #in the tens column. I use a student-t distribution with a low df (specifically 3)
        #instead of a normal distribution to account for outliers (essentially as a form of
        #robust regression). The censoring comes in two parts: interval censoring to account
        #for rounding of observations (since the responses are all whole numbers), and 
        #censoring at the endpoints to account for the bounded nature of the responses (from
        #1 to 99).
        for (i in 1:n) {
            log_p[3] : real     #temp for mixture log probabilities
    
            for (k in 1:3) {
                mu_offset : real

                if (k == 1 && mu[i] > 10) {
                    mu_offset = -10
                }
                else if (k == 3 && mu[i] < 90) {
                    mu_offset = 10
                }
                else {
                    mu_offset = 0
                }

                log_p[k] = log_diff_exp(
                    student_t_lcdf(response[i] + 0.5  | nu, mu[i] + mu_offset, sigma[i]),
                    student_t_lcdf(response[i] - 0.5  | nu, mu[i] + mu_offset, sigma[i])
                ) - log_diff_exp(
                    student_t_lcdf(99.5  | nu, mu[i] + mu_offset, sigma[i]),
                    student_t_lcdf(0.5  | nu, mu[i] + mu_offset, sigma[i])
                )

                #theta_k[k] is the probability for this portion of the mixture distribution
                log_p[k] = log_p[k] + log(theta_k[k])
            }
            target %+=% log_sum_exp(log_p)
        }
    },
    generated_quantities = {
        mu_offset : vector[n]
        k : vector[n]

        for (i in 1:n) {
            k[i] <- categorical_rng(theta_k)
            if (k[i] == 1 && mu[i] > 10) {
                mu_offset[i] = -10
            }
            else if (k[i] == 3 && mu[i] < 90) {
                mu_offset[i] = 10
            }
            else {
                mu_offset[i] = 0
            }
        }
    }
)
m.stan = stan_model(model_code = as.character(m.meta))
```

```{r}
fit.v = vb(m.stan, data = compose_data(filter(sv, vis == "squarepie")), algorithm="meanfield")
```

```{r}
fit = sampling(m.stan, data = compose_data(filter(sv, vis == "squarepie"))
    # , control = list(adapt_delta = 0.9)
    # , iter = 2000 * 6
    # , thin = 6
    )
```

```{r}
summary(fit, pars=c("mu_intercept","mu_b_p","sigma_intercept","sigma_v_participant_sd","nu","theta"))
precis(fit, pars=c("mu_intercept","mu_b_p","theta","sigma_intercept","sigma_v_participant_sd"))
plot(fit, pars=c("mu_intercept","mu_b_p","theta_k2","sigma_intercept","sigma_v_participant_sd"))
```

```{r}
gen.trun(c(0.5,99.5), family="TF", type="both")

samples = tidy_samples(fit, c(
    mu_intercept, mu_b_p, 
    sigma_intercept,
    sigma_v_participant_sd,
    nu
    )[]) %>%
    inner_join(
        tidy_samples(fit, theta_k[..]),
        by = ".sample"
    )


fitted = 
    tidy_samples(fit, c(mu, mu_offset, sigma)[i]) %>%
    inner_join(tidy_samples(fit, nu[])) %>%
    mutate(
        predicted = rTFtr(n(), mu + mu_offset, sigma, nu),
        predicted = round(predicted)
    ) %>%
    inner_join(mutate(filter(sv, vis=="squarepie"), i = 1:n()))

residuals = fitted %>%
    group_by(i, p, response, participant) %>%
    summarise(
        #randomized quantile residuals
        p_resid_max = mean(predicted < response + 0.5),
        p_resid_max = ifelse(p_resid_max == 0, 1/n(), p_resid_max),
        p_resid_min = mean(predicted < response - 0.5),
        p_resid_min = ifelse(p_resid_min == 1, (n() - 1)/n(), p_resid_min),
        p_resid = runif(1, p_resid_min, p_resid_max),
        z_resid = qnorm(p_resid),
        median_predicted = median(predicted),
        predicted = sample(predicted, 1),
        mu = mean(mu)
    ) %>%
    ungroup() %>%
    mutate(
        # z_resid = z_resid / sd(z_resid[is.finite(z_resid)]),
        residual = predicted - response
    )
```

```{r}
residuals %>% flatworm(z_resid, participant, ylim=NA) + facet_wrap(~.cuts)
```

```{r}
residuals %>% flatworm(z_resid)
```

```{r}
residuals %>% flatworm(z_resid, p)
```

```{r}
pairs(fit, pars=c("mu_intercept","mu_b_p","theta","sigma_intercept","sigma_v_participant_sd"))
pairs(fit, pars=c("sigma_intercept","sigma_v_participant_sd","sigma_v_participant[28]","sigma_v_participant[8]","sigma_v_participant[34]"))
pairs(fit, pars=c("sigma_intercept","sigma_v_participant[1]","sigma_v_participant[2]","sigma_v_participant[3]"))
```

```{r}
predictions = 
    expand.grid(
        p = 1:99
    ) %>%
    group_by(p) %>%
    do({
        cbind(., samples) %>%
            group_by(.sample) %>%
            mutate(
                k = rcategorical(1, c(theta_k1, theta_k2, theta_k3))
            ) %>%
            ungroup() %>%
            mutate(
                sigma_v_participant = rnorm(n(), 0, sigma_v_participant_sd),
                sigma = exp(sigma_intercept + sigma_v_participant),
                mu = mu_intercept + mu_b_p * (p - 50) + 50,
                mu_offset = 
                    ifelse(k == 1 & mu > 10, -10,
                    ifelse(k == 3 & mu < 90, 10,
                        0)),
                predicted = rTFtr(n(), mu + mu_offset, sigma, nu),
                predicted = round(predicted)
            )
    }) %>%
    mutate(error = predicted - p)
```

```{r}
predictions %>%
    # filter(p %in% (1:10 * 10)) %>%
    # group_by(p) %>%
    filter(p == 50) %>%
    mutate(
        # mu = mu_intercept + mu_b_p * (p - 50) + 50,
        predicted = rTFtr(n(), mu, sigma, nu)
    ) %>%
    summarise(
        above = mean(predicted > p + 10),
        below = mean(predicted < p - 10),
        theta = 1 - above - below
    )
```


```{r}
sv %>%
    filter(vis == "squarepie") %>%
    ggplot(aes(x = response - p)) +
    stat_density(fill=NA, color="red") +
    stat_density(aes(x=round(predicted) - p), fill=NA, color="green", data=residuals, adjust=1)
```

```{r}
fitted10 = fitted %>%
    filter(.sample %in% 15:20)
sv %>%
    filter(vis == "squarepie") %>%
    ggplot(aes(x = response - p)) +
    stat_density(fill=NA, color="black", size=1.5) +
    stat_density(aes(x=round(predicted) - p, group = .sample), fill=NA, color="blue", data=fitted10, adjust=1) +
    coord_cartesian(xlim=c(-20,20))
```

```{r}
predictions %>%
    select(-mu) %>% rename(mu = predicted) %>%
    mean_qi(mu) %>%
    ggplot(aes(x = p, y = mu)) +
    geom_point(aes(y = response), data=filter(sv, vis=="squarepie"), alpha=0.1) +
    geom_ribbon(aes(ymin = mu.lower, ymax = mu.upper), alpha = 0.25) +
    geom_abline(intercept = 0, slope = 1, color="gray50") +
    geom_line(color="red")
```

```{r}
predictions %>%
    filter(.sample %in% 1:1000) %>%
    ggplot(aes(x = p, y = round(predicted))) +
    geom_point(alpha=0.05)
```

```{r}
residuals %>%
    ggplot(aes(x = p, y = predicted)) +
    geom_point(alpha=0.1)
```
